---
title: "R - Mean, Median, Mode"
author: "MOHAMMAD SHADAN"
date: "January 16, 2017"
output: 
  html_document: 
    keep_md: yes
---


### Measures of Central Tendency

1. Mean   : Average   
2. Median : Middle Value   
3. Mode   : Most Often    


```{r}
x <- c(8,2,7,1,2,9,8,2,10,9)

#Exploratory Data Analysis
hist(x)
boxplot(x)
```

####Mean : The mean is the average of the numbers
```{r}
sum(x)/length(x)

#?mean
#Function in base R
mean(x)
```

####Median : the middle number given the numbers are in order (sorted)
```{r}
sort(x)

#?median
median(x)
```

####Mode : The number which appears most often in a set of numbers
```{r}
#There is no function in base R to find mode of set of numbers
x <- c(8,2,7,1,2,9,8,2,10,9)

#Function to find Mode
x
#?table
y <- table(x)
y
names(y)[which(y==max(y))]

#or in single line
names(table(x))[table(x)==max(table(x))]

#Testing if there are two or more numbers with same frequncy
x <- c(8,2,7,1,2,9,8,2,10,9,8)
sort(x)
#Mode
names(table(x))[table(x)==max(table(x))]
```

###Mean, Median and Mode using `mtcars` dataset

```{r}
head(mtcars)
x <- mtcars$wt
```

#### Mean
```{r}
mean(x)
```

#### Median
```{r}
median(x)
```

####Mode
```{r}

y <- table(x)
y
names(y)[which(y==max(y))]

#or
names(table(x))[table(x)==max(table(x))]
```

####Mean, Median and Mode using `airquality` dataset  

I am using `airquality` dataset because it has missing values

```{r}
#Summary Statistics
dim(airquality)
names(airquality)
str(airquality)
head(airquality)

#Column names with missing Values
names(airquality)[colSums(is.na(airquality)) > 0]

airquality$Ozone
airquality$Solar.R

x <- airquality$Solar.R
table(is.na(x))
x
```
####Mean
```{r}
mean(x)
?mean
mean(x, na.rm = TRUE)
```

####Median
```{r}
median(x)

median(x, na.rm = TRUE)
```

####Mode
We will not have issue of removing NA for finding Mode 
```{r}
sort(table(x))
names(table(x))[table(x)==max(table(x))]     #x<- airquality$Solar.R

# sort(table(x))
# sort(table(x, useNA = "always"))
```

####Summary fuctions

- summary()    #Base R
- describe()   #Package `psych`

```{r}
summary(mtcars)
summary(airquality)
```

```{r}
#install.packages("psych")
library(psych)
describe(mtcars)
describe(airquality)
```