---
title: "R Base Plotting"
author: "MOHAMMAD SHADAN"
date: "January 13, 2017"
output: 
  html_document: 
    keep_md: yes
---
Basic Video about the base plotting in R
Length of Video : 5 min

To do any kind of plotting the first thing you need is data.    
R comes pre-loaded with some data sets for you to practise    

Now how would you know which all data sets you can access without loading it externally ?    


```{r}
data()
# install.packages("datasets")
# library(datasets)
```

I will be using `mtcars` dataset.    
Load the dataset

```{r}
data(mtcars)
```

Summary Statistics
```{r}
dim(mtcars)      #dimension
str(mtcars)      #Internal STRucture
summary(mtcars)  #Summary
names(mtcars)    #Column Names
rownames(mtcars) #Row Names
dimnames(mtcars) #Row and Column Names together
head(mtcars)     #Top 6 entries
tail(mtcars)     #Last 6 entries
```

Shortcuts

- Ctrl + Enter : Runs the current line
- Ctrl + L     : Clears the Console
- Ctrl + Shift + C : Commenting a line or mutliple lines

###Base Plot   
1. Generic X-Y Plotting (Scatter, Line), plot()
2. Histogram, hist()
3. Barplot, barplot()
4. Boxplot, boxplot()

####Generic X-Y Plotting

```{r}
head(mtcars)
plot(mtcars$wt)
plot(sort(mtcars$wt))
plot(sort(mtcars$wt, decreasing = TRUE))
# ?plot
# ?sort
plot(mtcars$wt)
```

Different parameters in plot()

- `type` : type of plot (scatter or line etc.)

- `pch`: the plotting symbol (default is open circle)
- `cex`: symbol size in the plot, default is cex=1

- `lty`: the line type (default is solid line), can be dashed, dotted, etc.
- `lwd`: the line width, specified as an integer multiple. lwd=1 is default and 2 is twice as wide.

- `col`: the plotting color, specified as a number, string, or hex code; the colors() function gives  you a vector of colors by name

- `xlab`: character string for the x-axis label
- `ylab`: character string for the y-axis label
- `main`: character string for the main label (overall title for the plot)
- `sub` : a sub title for the plot

First parameter we look into is "type="    

- `p`	    points
- `l`	    lines
- `o`    overplotted points and lines
- `b,c` 	  points (empty if "c") joined by lines
- `s,S` 	  stair steps
- `h`	    histogram-like vertical lines
- `n`	    does not produce any points or lines


```{r}
plot(mtcars$wt, type="p")     # points (default)
plot(mtcars$wt, type="l")     # lines
plot(mtcars$wt, type="o")     # points and lines together

plot(mtcars$wt, type="b")     # for both points and lines
plot(mtcars$wt, type="c")     # for the lines part alone of "b",
plot(mtcars$wt, type="s")     # stair steps
plot(mtcars$wt, type="S")     # stair steps
plot(mtcars$wt, type="h")     # histogram like vertical lines
plot(mtcars$wt, type="n")     # does not produce any points (???)
```

Parameter pch and cex. This is for points i.e. default or type="p"
```{r}
plot(mtcars$wt)
plot(mtcars$wt, type="p", pch=1, cex=1)     # default
plot(mtcars$wt, type="p", pch=2, cex=1) 
plot(mtcars$wt, type="p", pch=2, cex=2) 
plot(mtcars$wt, type="p", pch=18, cex=3) 
plot(mtcars$wt, type="p", pch=18, cex=3, col="red")
plot(mtcars$wt, type="p", pch=18, cex=3, col="blue")
```

[Plotting Symbols](http://www.statmethods.net/advgraphs/parameters.html)
[Colors](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf)

Parameter lty and lwd. This is for points i.e. default or type="p"
```{r}
plot(mtcars$wt, type="l")
plot(mtcars$wt, type="l", lty=1, lwd=1)     # points
plot(mtcars$wt, type="l", lty=2, lwd=3)     # points
plot(mtcars$wt, type="l", lty=3, lwd=3)
colors()
```
[Line Type](http://www.statmethods.net/advgraphs/parameters.html)

###Parameters xlab, ylab, main

```{r}
#Default Values

plot(mtcars$wt)

plot(mtcars$wt,
     type="p", #points
     pch=1,    #point shape
     cex=1,    #point size
     lty=1,    #line type
     lwd=1,    #line width
     ylab = "Weight",
     xlab = "Weight Index",
     main = "Distribution of Car Weight")
```

Let's Try all together with different parameters
```{r}
plot(mtcars$wt,
     type="o", #points
     pch=18,    #point shape
     cex=2,    #point size
     lty=5,    #line type
     lwd=2,    #line width
     col="red",
     xlab = "Weight",
     ylab = "Weight Index",
     main = "Distribution of Car Weight",
     sub = "Sub Title")
```
####Scatter Plot with two variables
```{r}
plot(x=mtcars$wt, y=mtcars$mpg)
```
